Chasing \rainbows{} and ocean glint:\\ Inner working angle constraints for the Habitable Worlds Observatory

### 2023-05-02 src/tex/ms.tex
[showyourwork] overleaf sync

%--------------------------------------------------------------------------
% Figure 4
%--------------------------------------------------------------------------
\begin{figure*}
    \centering
    \includegraphics{figures/figure-4-scatterplot.pdf}
    \script{create-figure-4-scatterplot.py}
    \caption{
        Scatter plot for the target star sample, showing stellar effective 
        temperatures and distances.
        %
        The circle sizes represent the angular separation between 
        the star and its planet (in mas) as presented in the target list.
        %
        The colours show $\Delta \varphi$ assuming circular, edge-on orbits
        at a semi-major axis $a$ corresponding to an Earth-like 
        instellation and an IWA of \qty{62}{\mas}.
        %
        Additionally, the colour bar indicates the optical phenomena that 
        can in principle be detected: the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
        %
        For example, dark blue circles indicate planets on which most phenomena would be observable because planets that can be observed at the rainbow angle can also be observed at angles exhibiting Rayleigh scattering.
    }
    \label{fig:scatterplot}
\end{figure*}
%--------------------------------------------------------------------------

@@

The NASA Exoplanet Exploration Program’s Mission Star List for \hwo{} 
includes for every star an estimated contrast for an Earth-twin in 
the habitable zone.
%
% The cited paper has all the equations that lead to this equation:
The contrast is given by \citep[see, e.g.,][]{2023A&A...671A.165M}
\begin{equation}
    C(\alpha) 
    = F_p(\alpha) / F_* 
    = A_\mathrm{G} \cdot \phi(\alpha) \cdot \left( R_p / a(\alpha) \right)^2 \,,
\label{eq:contrast}
\end{equation}
where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the integral phase function at scattering angle $\alpha$
(normalized to \num{1.0} at full phase), and $a$ is the distance between the planet and its host star.
\todo{here, $\alpha$ is used for the scattering angle ... and is the 
'integral phase function' the same as the 'scattering phase function'?}
%
The assumptions that went into the contrast estimations for \hwo{}
are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the scattering phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
%
These calculations give an order-of-magnitude calculation for the 
contrast of Earth-like planets around nearby stars.
%
Therefore, in this work, we will assume that the calculated contrasts 
in the list correspond to the reflected flux at quadrature and normalize the scattering phase functions to this point.
%
We do not aim to accurately calculate the absolute scaling of the contrast, as this depends on many factors and would have required some work for the star list to be redone.
%

We use the scattering phase functions from \citet{treesandstam2019} 
for an Earth-like planet with an ocean surface with a wind-speed of $\qty{7}{\meter\per\second}$ at $\qty{670}{\nano\meter}$ and covered by patchy clouds.
%
The $\qty{670}{\nano\meter}$ is close to the centre of the Rc band for Vega ($\qty{642}{\nano\meter}$). 
%
We normalize the phase function at quadrature and multiply it with the 
contrast.
%
In addition, we map the orbital phase for circular orbits to the phase angle and the on-sky separation given an inclination of \num{90} degrees.
%
Together, they give the reflected light contrast as a function of separation for a full orbit.
%
We multiply this flux with the degree of polarization and retrieve the polarized contrast.
%
The resulting contrast curves are shown in \cref{fig:contrasts} for 
three of the closest target stars: the G~dwarf $\alpha$~Cen~A, the K~dwarf 
$\epsilon$~Eri, and the M~dwarf Lalande~21185.

>>>>

%--------------------------------------------------------------------------
% Figure 4
%--------------------------------------------------------------------------
\begin{figure*}
    \centering
    \includegraphics{figures/figure-4-scatterplot.pdf}
    \script{create-scatterplot.py}
    \caption{
        Scatter plot for the target star sample, showing stellar effective 
        temperatures and distances.
        %
        The circle sizes represent the angular separation between 
        the star and its planet (in mas) as presented in the target list.
        %
        The colours show $\Delta \varphi$ assuming circular, edge-on orbits
        at a semi-major axis $a$ corresponding to an Earth-like 
        instellation and an IWA of \qty{62}{\mas}.
        %
        Additionally, the colour bar indicates the optical phenomena that 
        can in principle be detected: the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
        %
        For example, dark blue circles indicate planets on which most phenomena would be observable because planets that can be observed at the rainbow angle can also be observed at angles exhibiting Rayleigh scattering.
    }
    \label{fig:scatterplot}
\end{figure*}
%--------------------------------------------------------------------------

@@

The NASA Exoplanet Exploration Program’s Mission Star List for \hwo{} 
includes for every star an estimated contrast for an Earth-twin in 
the habitable zone.
%
% The cited paper has all the equations that lead to this equation:
The contrast is given by \citep[see, e.g.,][]{2023A&A...671A.165M}
\begin{equation}
    C(\alpha) 
    = F_p(\alpha) / F_* 
    = A_\mathrm{G} \cdot \phi(\alpha) \cdot \left( R_p / a(\alpha) \right)^2 \,,
\label{eq:contrast}
\end{equation}
where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$
(normalized to \num{1.0} at full phase, thus at $\alpha=0^\circ$), 
and $a$ is the distance between the planet and its host star.
%
The assumptions that went into the contrast estimations for \hwo{}
are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
%
These calculations give an order-of-magnitude calculation for the 
contrast of Earth-like planets around nearby stars.
%
Therefore, in this work, we will assume that the calculated contrasts 
in the list correspond to the reflected flux at quadrature and normalize the phase functions to this point.
%
We do not aim to accurately calculate the absolute scaling of the contrast, as this depends on many factors and would have required some work for the star list to be redone.
%

We use the phase functions from \citet{treesandstam2019} 
for an Earth-like planet with an ocean surface with a wind-speed of $\qty{7}{\meter\per\second}$ at $\qty{670}{\nano\meter}$ and covered by patchy clouds.
%
The $\qty{670}{\nano\meter}$ is close to the centre of the Rc band for Vega ($\qty{642}{\nano\meter}$). 
%
We normalize the phase function at quadrature and multiply it with the 
contrast.
%
In addition, we map the orbital phase for circular orbits to the phase angle and the on-sky separation given an inclination of \num{90} degrees.
%
Together, they give the reflected light contrast as a function of separation for a full orbit.
%
We multiply this flux with the degree of polarization and retrieve the polarized contrast.
%
The resulting contrast curves are shown in \cref{fig:contrasts} for 
three of the closest target stars: the G~dwarf $\alpha$~Cen~A, the K~dwarf 
$\epsilon$~Eri, and the M~dwarf Lalande~21185.
